---
title: "Mega Awesome Beer Corp. (MABco) Case Study"
author: "Data Busters Inc. (K Henley, G Lane, J McDaniel)"
date: "October 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Mega Awesome Beer Corp. (MABco), has commissioned Data Busters, Inc.(DBI) to examine data sets they have provided, with the purpose of answering some key questions and findings that should be derived directly from the data provided.  These are the results. 

Two data sets were provided by MABco and used by DBI to provide the findings below.

```{r, echo=TRUE}
beerData <- read.csv2(url("https://raw.githubusercontent.com/GriffinLane/DataBustersCaseStudy/master/Beers.csv"), header=TRUE, sep=",", col.name=c("Name", "BeerID", "ABV", "IBU", "BreweryID", "Style", "Ounces"))

breweriesData <- read.csv2(url("https://raw.githubusercontent.com/GriffinLane/DataBustersCaseStudy/master/Breweries.csv"), header=TRUE, sep=",", col.name=c("BreweryID", "BreweryName", "City", "State"))
```

Below is the summary for both data sets.  As you can see the data provided is organized
into 2410 types of beer, and between the two sets of data we have 11 measurable variables 
to work with.  We will walk you through the results below, and provide you with the
background necessary to understand the context behind our findings.
```{r, echo=TRUE}
summary (beerData)
str(beerData)
summary (breweriesData)
str(breweriesData)
```

In order to take full advantage of the data provided we have opted to merge the two files.
The new merged data set contains ALL the measurable variables in one file called "merged"

```{r,  echo=TRUE}
merged <- merge(beerData, breweriesData, by = "BreweryID", all=FALSE)
```

To further organize our data we have taken a look at the first and last few entries within
our newly merged data set.  This will allow us to a) ensure we have a better understanding of
the new data file we are working with, and b) help to ensure all data was properly pulled 
into the file as expected.  Upon inspection of the 'merge' file we see no unexpected issues.

```{r, echo=TRUE}
head(merged)
tail(merged)
```

As expected, we do encounter some missing data that are currently labeled as NA. However, the ABV column is currently stored in the data set as an ordered factor. In order to rectify this we must first convert the data to a numeric type. This wil allow us to obtain the NAs rather than treating them as an element of the ordered factor.

```{r, echo=TRUE}
##First we must fix the numeric columns incorrectly listed as ordered factors
##A direct conversion to a numeric will instead return the rank of the factor
##Converting to a matrix first avoids this issue
merged$ABV <- as.numeric(as.matrix(merged$ABV))
merged$Ounces <- as.numeric(merged$Ounces)



colSums(is.na(merged))
```

In order to provide a neater representation on a per state basis, we have sorted the data frame by state and city. This will allow for state subsections of the data to be neatly ordered by city in the future.

```{r, echo=TRUE}
##This line will order our merged data by state, then by city and store it in a new data frame called simply "beer""
beer <- merged[order(merged$State, merged$City),] 
```

Given the merged results DBI was able to confirm that MABco's presence in each U.S.
state is as follows:

```{r, echo=TRUE}
##The Unique() function is used to ensure the same brewery is not recorded multiple times for each state
table(beer$State[unique(beer$BreweryID)])
```

DBI was also tasked with providing a better understanding of MABco's beer offerings in
terms of alcohol content (ABV).  We determined that the highest median ABV is 0.092, and
was reported in the states of Georgia, Florida, and Idaho when computing the medians for all beer offerings in each state respectively.

We have also provided median IBU values for each state for reference in the table below.
```{r, echo=TRUE}

##Create columns for both measurements to allow for quick lookups in the future on a per tuple basis
beer$StateABVMedian <- sapply(beer$ABV[beer$State], median)
beer$StateIBUMedian <- sapply(beer$IBU[beer$State], median)

##Create a list of every state included and store it in a variable
allStates <- unique(beer$State[beer$BreweryID])

##Use that list to index the new columns and create a data frame to display the findings
mediansByState <- data.frame(State=allStates, ABV_median = beer$StateABVMedian[allStates], IBU_median = beer$StateIBUMedian[allStates])

##Display the ne data frame that holds the median data
mediansByState
``` 

To help MABco better understand these results we've plotted them graphically.
```{r, echo=TRUE}
library(ggplot2)
##Create a plot with the staes along the x-axis
p <- ggplot2::ggplot(mediansByState, aes(x = State))

##Create and add the y-axis. Also, give it a low alpha value and change it's fill color
p <- p + geom_bar(aes(y = IBU_median, color = "IBU"), stat="identity", alpha = .5 , fill="blue")

##Create and add the second y-axis. Also, give it a low alpha value and change it's fill color
##Note that the extra 100 multiplier is to turn the scale of those data into a percentage scale
p <- p + geom_bar(aes(y = ABV_median*100*6, color = "ABV"), alpha = .5 , fill="red", stat="identity")

##Create the right side second axis and give it the inverse scale used to transform the data
p <- p + scale_y_continuous(sec.axis = sec_axis(~./6, name = "ABV")) 

##Display the ggplot and marvel at its splendor
p
```


Upon review of the results we derived from looking at the median alcohol content (ABV),
as well as the median bitterness levels (IBU) we are able to confirm that the states with
beer offerings with the highest median ABV are Georgia, Florida, and Idaho (with median ABVs of 0.092 respectively) as previously confirmed.  

**However the beer with the highest ABV is Lee Hill Series Vol. 5 - 
Belgian Style Quadrupel Ale (0.128 ABV).  This beer is sold in COLORADO.**
```{r, echo=TRUE}
##This line indexes the state by grabbing the index of the tuple with the highest ABV using the where() function
beer$State[which.max(beer$ABV)]
```

We are further able to confirm that the states with the highest median bitter beers as measured in IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ respectively, with each state returning
a median IBU score of 135.  
**However, the individual state returning the highest IBU (138)is OREGON (Bitter Bitch Imperial IPA)**

```{r, echo=TRUE}
##This line indexes the state by grabbing the index of the tuple with the highest IBU rating using the where() function
beer$State[which.max(beer$IBU)]
```

In order to proved a clearer picture of the Alcohol By Volume, its summary will be displayed below
```{r}
##In order to display as a human readable percentage, the ABV will be scaled up by a factor of 100.
##Also, please note that there are still 62 NA values that could not be factored into the statistics below
summary(beer$ABV * 100) 
```



*What is the relationship of Alcohol content and Bitterness for MABco's beer offerings?*
The graph below is a scatterplot that examines the relationship between ABV and IBU for each
beer that MABco has provided us data on.  As you can see there is a direct linear correlation 
betwen the bitterness of the beer and it's corresponding alcohol content.


```{r, echo=TRUE}
## Scatterplot examining relationship between bitterness and alcohol content ##
## ABV vs. IBU 
library(car)
scatterplot(ABV ~ IBU, data=merged, 
  	xlab="Alcohol Content in ABV", ylab="Bitterness in IBU", 
   main="Relationship between ABV and IBU", 
   labels=row.names(merged))
```



##WE NEED TO SUMMARIZE OUR FINDINGS HERE##


