---
title: "Mega Awesome Beer Corp. (MABco) Case Study"
author: "Data Busters Inc. (K Henley, G Lane, J McDaniel)"
date: "October 19, 2017"
output: html_document
---

Mega Awesome Beer Co. (MABco), has commissioned Data Busters Incorporated (DBI), to examine data sets to answer key questions and present findings derived from the data provided.  These are the results. 

Two data sets were provided to DBI by MABco. One data set presented background on key craft beer offerings currently in production in the United States. The second consisted of data from where each beer was produced. The raw data was then ran into DBI's system
for analysis.

```{r}
beerData <- read.csv2(url("https://raw.githubusercontent.com/GriffinLane/DataBustersCaseStudy/master/Beers.csv"),
header=TRUE, sep=",", col.name=c("Name", "BeerID", "ABV", "IBU",
"BreweryID", "Style", "Ounces"), stringsAsFactors = TRUE)

breweriesData <-
read.csv2(url("https://raw.githubusercontent.com/GriffinLane/DataBustersCaseStudy/master/Breweries.csv"),
header=TRUE, sep=",", col.name=c("BreweryID", "BreweryName", "City",
"State"), stringsAsFactors = TRUE)
```

Below is the summary for both data sets. The data provided is organized
into 2410 types of beer. Between the two sets of data there are 10 variables.  
DBI will walk you through the results below, provide you with the
background necessary to understand the context behind the findings, and hopefully answer any questions that may arise.
```{r}
summary (beerData)
str(beerData)
summary (breweriesData)
str(breweriesData)
```

In order to take full advantage of the data provided we have opted to merge the two files.
The new merged data set contains ALL the measurable variables in one file called "merged",
but will now allow us to cross reference the beer and brewery data for a more complete analysis.
```{r}
merged <- merge(beerData, breweriesData, by = "BreweryID", all=TRUE)
```

To further organize our data, we have taken a look at the first and last few entries within
our newly merged data set. This will allow us to a) ensure we have a better understanding of
the new data file we are working with, and b) help to ensure all data was properly pulled 
into the file as expected.  Upon inspection of the first 6 and last 6 entries within the 'merged' file we see no unexpected issues.
```{r}
head(merged)
tail(merged)
```

As expected, we did encounter some instances of missing data after the merge. Specifically, there were 1067 instances in total of missing datum that were not measured/recorded and are marked as "NA".  There were 62 instances in the ABV column and 1005 in the IBU column.  
```{r}
##First we must fix the numeric columns incorrectly listed as ordered factors
##A direct conversion to a numeric will instead return the rank of the factor
##Converting to a matrix first avoids this issue
merged$ABV <- as.numeric(as.matrix(merged$ABV))
merged$IBU <- as.integer(merged$IBU)
merged$Ounces <- as.numeric(merged$Ounces)

##This line will return the sum of all NAs in each column
colSums(is.na(merged))
```

In order to provide a neater representation on a per state basis, we have sorted the data frame by state and city. This will allow for state subsections of the data to be neatly ordered by city in the future.  In order to further understand competitor brewery presence in each U.S. state, we have provided a summary table below:
```{r}
##This line will order our merged data by state, then by city and store it in a new data frame called simply "beer""
beer <- merged[order(merged$State, merged$City),] 

##Table is a useful function getting the frequencies of something
table(beer$State[!duplicated(beer$BreweryID)])
```

DBI was also tasked with providing a better understanding of available beer offerings in
terms of alcohol content (ABV).  We determined that the highest median ABV is 0.092, and
was reported in the states of Georgia, Florida, and Idaho when computing the medians for all beer offerings in each state respectively.

We have also provided median IBU values for each state for reference in the table below.
```{r}
##Create columns for both measurements to allow for quick lookups in the future
StateABVMedian <- aggregate(beer$ABV, by=list(beer$State),FUN=median, na.rm = TRUE)[,2]
StateIBUMedian <- aggregate(beer$IBU, by=list(beer$State),FUN=median, na.rm = TRUE)[,2]

##Create a list of every state included and store it in a variable
allStates <- beer$State[!duplicated(beer$State)]

##Use that list to index the new columns and create a data frame to display the findings
##The indexing was probably not necessary since the data are in the same order but it will be included anywyay
mediansByState <- data.frame(State=allStates, ABV_median = StateABVMedian[allStates], IBU_median = StateIBUMedian[allStates])

##Display the ne data frame that holds the median data
mediansByState
```

To help MABco better understand these results, they have been plotted graphically for reference.
```{r}
library(ggplot2)
##Create a plot with the staes along the x-axis
p <- ggplot2::ggplot(mediansByState, aes(x = State))

##Create and add the y-axis. Also, give it a low alpha value and change it's fill color
p <- p + geom_bar(aes(y = IBU_median, color = "IBU"), stat="identity", alpha = .5 , fill="blue")

##Create and add the second y-axis. Also, give it a low alpha value and change it's fill color
##Note that the extra 100 multiplier is to turn the scale of those data into a percentage scale
p <- p + geom_bar(aes(y = ABV_median*100*6, color = "ABV"), alpha = .5 , fill="Green", stat="identity")

##Create the right side second axis and give it the inverse scale used to transform the data
p <- p + scale_y_continuous(sec.axis = sec_axis(~./6, name = "ABV")) 

##This line orients the x-axis vertically and moves the labels slightly closer to the plot
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1))

##Display the ggplot and marvel at its splendor
p
```

Upon review of the results of the median alcohol content (ABV), and the median bitterness levels (IBU), we are able to confirm that the states offering beer with the highest median ABV are Georgia, Florida, and Idaho (with median ABVs of 0.092 respectively).  

**However the beer with the highest ABV is Lee Hill Series Vol. 5 - 
Belgian Style Quadrupel Ale with an ABV of 12.8% (or .0128 as it appears in the datset).  This beer is sold in COLORADO.**
```{r, echo=TRUE}
beer$State[which.max(beer$ABV)]
```

We are further able to confirm that the states with the highest median bitter beers as measured in IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ with each state returning a median IBU score of 135.  
**However, the individual state returning the highest IBU (138)is OREGON (Bitter Bitch Imperial IPA)**

```{r, echo=TRUE}
beer$State[which.max(beer$IBU)]
```

In order to provide a clearer picture of the Alcohol By Volume, see the summary below.
```{r}
##In order to display as a human readable percentage, the ABV will be scaled up by a factor of 100.
##Also, please note that there are still 62 NA values that could not be factored into the statistics below
summary(beer$ABV * 100) 
```

*What is the relationship of Alcohol content and Bitterness for the available beer offerings?*
The graph below is a scatterplot that examines the relationship between ABV and IBU for each
beer MABco offered data on. There is a direct linear correlation between the bitterness of the beer and the corresponding alcohol content. As the alcohol content increases, so does the bitterness.


```{r, echo=TRUE}
## Scatterplot examining relationship between bitterness and alcohol content ##
## ABV vs. IBU 
library(car)
scatterplot(ABV ~ IBU, data=merged, 
  	xlab="Alcohol Content in ABV", ylab="Bitterness in IBU", 
   main="Relationship between ABV and IBU", 
   labels=row.names(merged))
```



DBI has provided MABco with a look at brewery presence (by U.S. State) for key beer offerings, and the results show the top five to be Texas (255), Colorado (253), Indiana (199), Oregon (188), and New York (168).  We've also provided data regarding the median Alcohol by volume (ABV) by each state.  Our findings determined that the states with the highest median (ABV) are Georgia, Florida, and Idaho, each recording a median of 0.092 respectively.  We also called out that although these states recorded the highest median values, the state with highest overall ABV from a single beer offering was found outside of this population.  Colorado sells the beer with the overall highest ABV offered (Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale (0.128 ABV)).  When examining all beer offerings the overall average (ABV) was found to be 0.05977.  Further, we provided data regarding the bitterness of the beer offerings as measured in (IBU), and reported our findings by state. DBI determined that the states with the highest median bitter beers as measured in IBU's are NY, OH, NC, OR, PA, NM, MT, SC, and NJ respectively, with each state returning a median IBU score of (135 IBU).  When measuring (IBU) overall maximum's we found that the individual state returning the highest IBU (138) was Oregon (Bitter Bitch Imperial IPA).  Please note the state with the maximum bitterness value does reside in the highest median population we examined (unlike the ABV findings).  Lastly, we provided a graphical representation outlining the relationship between the Alcohol by Volume (ABV) and the bitterness (IBU) of each of the beers in our data set.  The graph clearly outlined a positive linear relationship between the variables.  In general, as the ABV increased, so did the bitterness levels as recorded by IBU.  Further evaluation can be conducted here with MABco's direction if more detailed analysis is needed regarding the noted correlation.

This concludes our analysis.


