---
title: "Case Study 2_Henley_Lane_McDaniel"
author: "Kris Henley"
date: "November 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Initial exploratory data analysis.
*Intent:  uncover top 3 factors contributing to attition
*Intent:  identify specific trends that emerge regarding specific job roles


##look at data by job role and attrition
```{r, echo=True}
library(plyr)
AttritionbyRole <- CaseStudy2_data[c("Attrition", "JobRole")]
AttritionbyRole
```

```{r, echo=True}
##ggplot barchart
library(ggplot2)

##Barplot showing count of attrition by job roles
p <- ggplot2::ggplot(AttritionbyRole, aes(y = Attrition), fill=JobRole)
p <- p + geom_bar(aes(x = JobRole, color = JobRole), stat="identity")
## Flip the coordinates to better read the States Names
p <- p + coord_flip()
## Add a main title and center
p <- p + labs(title="Count of Attrition Instances by Job Role")
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(text = element_text(size=7))
##Change axis labels to something more informative
p <- p + labs(y="Attrition Count")
p <- p + labs(x="JObRole")  
p

```

*Top 3 Attrition counts by Role: 
1) Sales Executive
2) Research Scientist
3) Laboratory Technician

*Lowest 3 Attrition counts by Role:
1) Human Resources
2) Research Director
3) Manager


##creating subsets for each category

```{r, echo=TRUE}
#Top
TopAttrition <- subset(CaseStudy2_data, JobRole %in% c("Sales Executive","Research Scientist", "Laboratory Technician"))
SalesExecutive <- subset(CaseStudy2_data, JobRole=="Sales Executive")
ResearchScientist <- subset(CaseStudy2_data, JobRole=="Research Scientist")
LaboratoryTechnician <- subset(CaseStudy2_data, JobRole=="Laboratory Technician")

#Bottom
BottomAttrition <- subset(CaseStudy2_data, JobRole %in% c("Human Resources","Research Director", "Manager"))
HumanResources <- subset(CaseStudy2_data, JobRole=="Human Resources")
ResearchDirector <- subset(CaseStudy2_data, JobRole=="Research Director")
Manager <- subset(CaseStudy2_data, JobRole=="Manager")
```

##creating mean variables for our numeric data

```{r, echo=TRUE}
MeanAge <- mean(TopAttrition$Age)
MeanEnvironmentSatisfaction <- mean(TopAttrition$EnvironmentSatisfaction)
MeanHourlyRate <-mean(TopAttrition$HourlyRate)
MeanJobInvolvement <-mean(TopAttrition$JobInvolvement)
MeanJobLevel <-mean(TopAttrition$JobLevel)
MeanJobSatisfaction <-mean(TopAttrition$JobSatisfaction)
MeanMonthlyIncome <-mean(TopAttrition$MonthlyIncome)
MeanMonthlyRate <-mean(TopAttrition$MonthlyRate)
MeanNumCompaniesWorked <-mean(TopAttrition$NumCompaniesWorked)
MeanPercentSalaryHike <-mean(TopAttrition$PercentSalaryHike)
MeanPerformanceRating <-mean(TopAttrition$PerformanceRating)
MeanRelationshipSatisfaction <-mean(TopAttrition$RelationshipSatisfaction)
MeanStockOptionLevel <-mean(TopAttrition$StockOptionLevel)
MeanTotalWorkingYears <-mean(TopAttrition$TotalWorkingYears)
MeanTrainingTimesLastYear <-mean(TopAttrition$TrainingTimesLastYear)
MeanWorkLifeBalance <-mean(TopAttrition$WorkLifeBalance)
MeanYearsAtCompany <-mean(TopAttrition$YearsAtCompany)
MeanYearsInCurrentRole <-mean(TopAttrition$YearsInCurrentRole)
MeanYearsSinceLastPromotion <-mean(TopAttrition$YearsSinceLastPromotion)
MeanYearsWithCurrManager <-mean(TopAttrition$YearsWithCurrManager)

MeanAge2 <- mean(BottomAttrition$Age)
MeanEnvironmentSatisfaction2 <- mean(BottomAttrition$EnvironmentSatisfaction)
MeanHourlyRate2 <-mean(BottomAttrition$HourlyRate)
MeanJobInvolvement2 <-mean(BottomAttrition$JobInvolvement)
MeanJobLevel2 <-mean(BottomAttrition$JobLevel)
MeanJobSatisfaction2 <-mean(BottomAttrition$JobSatisfaction)
MeanMonthlyIncome2 <-mean(BottomAttrition$MonthlyIncome)
MeanMonthlyRate2 <-mean(BottomAttrition$MonthlyRate)
MeanNumCompaniesWorked2 <-mean(BottomAttrition$NumCompaniesWorked)
MeanPercentSalaryHike2 <-mean(BottomAttrition$PercentSalaryHike)
MeanPerformanceRating2 <-mean(BottomAttrition$PerformanceRating)
MeanRelationshipSatisfaction2 <-mean(BottomAttrition$RelationshipSatisfaction)
MeanStockOptionLevel2 <-mean(BottomAttrition$StockOptionLevel)
MeanTotalWorkingYears2 <-mean(BottomAttrition$TotalWorkingYears)
MeanTrainingTimesLastYear2 <-mean(BottomAttrition$TrainingTimesLastYear)
MeanWorkLifeBalance2 <-mean(BottomAttrition$WorkLifeBalance)
MeanYearsAtCompany2 <-mean(BottomAttrition$YearsAtCompany)
MeanYearsInCurrentRole2 <-mean(BottomAttrition$YearsInCurrentRole)
MeanYearsSinceLastPromotion2 <-mean(BottomAttrition$YearsSinceLastPromotion)
MeanYearsWithCurrManager2 <-mean(BottomAttrition$YearsWithCurrManager)

```

##creating new df with only numeric mean data from above for top attrition

```{r, echo=TRUE}
MeanTopAttrition = data.frame(TopAttrition$JobRole, MeanAge, MeanEnvironmentSatisfaction, MeanHourlyRate, MeanJobInvolvement, MeanJobLevel, MeanJobSatisfaction,MeanMonthlyIncome,
MeanMonthlyRate, MeanNumCompaniesWorked, MeanPercentSalaryHike, MeanPerformanceRating, MeanRelationshipSatisfaction, MeanStockOptionLevel, MeanTotalWorkingYears,
MeanTrainingTimesLastYear, MeanWorkLifeBalance, MeanYearsAtCompany, MeanYearsInCurrentRole, MeanYearsSinceLastPromotion, MeanYearsWithCurrManager)
MeanTopAttrition
```

##creating new df with only numeric mean data from above for bottom attrition

```{r, echo=TRUE}
MeanBottomAttrition = data.frame(BottomAttrition$JobRole, MeanAge2, MeanEnvironmentSatisfaction2, MeanHourlyRate2, MeanJobInvolvement2, MeanJobLevel2, MeanJobSatisfaction2,MeanMonthlyIncome2,
MeanMonthlyRate2, MeanNumCompaniesWorked2, MeanPercentSalaryHike2, MeanPerformanceRating2, MeanRelationshipSatisfaction2, MeanStockOptionLevel2, MeanTotalWorkingYears2,
MeanTrainingTimesLastYear2, MeanWorkLifeBalance2, MeanYearsAtCompany2, MeanYearsInCurrentRole2, MeanYearsSinceLastPromotion2, MeanYearsWithCurrManager2)
MeanBottomAttrition
```






































